---
title: "Debugging Worksheet"
format: html
editor: source
---

## Goal

Run the code. It crashes.

Use a **breakpoint** to find the bug.

------------------------------------------------------------------------

## Find the Error (Dissolution Code Example)

```{r}
# Preliminaries
library(tidyverse)
data <- read_csv("dissolution_data.csv")

# Standard curve parameters: A = slope * C (µg/mL) + intercept
std_intercept <- -0.0002667 # absorbance units (y-intercept)
```

1.  Try run the code.
2.  Insert `browser()` somewhere before the error occurs.
3.  Find the line that should be added to fix the bug.

```{r}
data <- data %>%
  mutate(
    conc_mg_ml = ((absorbance - std_intercept) / std_slope) * diluton / 1000
  )
```
------------------------------------------------------------------------

## Understanding Code (Dissolution Code Example)

```{r}
# Preliminaries
library(tidyverse)
data <- read_csv("dissolution_data.csv")

# Standard curve parameters: A = slope * C (µg/mL) + intercept
std_slope     <- 0.04478    # absorbance units per µg/mL
std_intercept <- -0.0002667 # absorbance units (y-intercept)
```
1. Clear the environment, the run the code block above.
2. Insert `browser()` within the loop in the code block below & run it.
3. Try out entering `n` to run the next line or `c`to run the code until the next `browser` in the Console to walk through the code.
4. See how variable values change in the Environment window.
5. Try understanding what the code does.

```{r}
small_data <- data[1:3, ]
small_data$conc_mg_ml <- NA_real_

# nrow(small_data) gives the number of rows in the table
for (i in 1:nrow(small_data)) {
  
  browser()
  
  absorbance_i <- small_data$absorbance[i]
  dilution_i   <- small_data$dilution[i]
  
  conc_i <- ((absorbance_i - std_intercept) / std_slope) * dilution_i / 1000
  
  small_data$conc_mg_ml[i] <- conc_i
}

small_data

```