---
title: "CodeClub3"
format: html
editor: visual
---

## Welcome! 👋
We will look at some trial data where a drug (Lisinopril) for hypertension treatment is tested. There are 3 groups of patients: one gets 10 mg of Lisinopril, one gets 20 mg of Lisinopril, and one gets a placebo without a drug. 

The effects (lowered blood pressure) is measured for every patient.


------------------------------------------------------------------------

## Step 1: Load packages

```{r}
# If you don't have tidyverse installed, run this once:
# install.packages("tidyverse")

library(tidyverse)
```

------------------------------------------------------------------------

## Step 2: Load data
Make sure your working directory is set correctly!

```{r}
library(readxl)

study_data <- read_excel("bp_trial.xlsx")

study_data
```

## Step 3: Plot the data

### Does a bar chart makes sense here? Not really!

```{r}
ggplot(study_data, aes(x=treatment, y=bp_reduction_mmHg)) + 
  geom_bar(stat = "identity")
```
### A scatterplot makes more sense!

```{r}
ggplot(study_data, aes(x = treatment, y = bp_reduction_mmHg)) +
  geom_point()
```
### Compare both!The current bar chart setting just sums up all of the values per treatment group.
```{r}
ggplot(study_data, aes(x=treatment, y=bp_reduction_mmHg)) + 
  geom_bar(stat = "identity") +
  geom_point()
```


### Add theme minimal
```{r}
ggplot(study_data, aes(x = treatment, y = bp_reduction_mmHg)) +
  geom_point() +
  theme_minimal()
```

### Color the dots
```{r}
ggplot(study_data, aes(x = treatment, y = bp_reduction_mmHg, color = treatment)) +
  geom_point() +
  theme_minimal()
```
### Change color map (might not be included in the live coding)
```{r}
# https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html
# install.packages("viridis")
library(viridis)

ggplot(study_data, aes(x = treatment, y = bp_reduction_mmHg, color = treatment)) +
  geom_point() +
  theme_minimal() +
  scale_color_viridis(discrete=TRUE)
```


### Add labels

```{r}
ggplot(study_data, aes(x = treatment, y = bp_reduction_mmHg, color = treatment)) +
  geom_point() +
  theme_minimal() +
  labs(
    title = "Blood Pressure Reduction by Treatment",
    x     = "Treatment",
    y     = "BP Reduction (mmHg)"
    )
```
### Remove the legend

```{r}
ggplot(study_data, aes(x = treatment, y = bp_reduction_mmHg, color = treatment)) +
  geom_point() +
  theme_minimal() +
  labs(
    title = "Blood Pressure Reduction by Treatment",
    x     = "Treatment",
    y     = "BP Reduction (mmHg)"
    ) +
  theme(legend.position = "none")
```
### Add some Jitter

```{r}
ggplot(study_data, aes(x = treatment, y = bp_reduction_mmHg, color = treatment)) +
  geom_jitter() +
  theme_minimal() +
  labs(
    title = "Blood Pressure Reduction by Treatment",
    x     = "Treatment",
    y     = "BP Reduction (mmHg)"
    ) +
  theme(legend.position = "none")
```
### Add some derived values

```{r}
ggplot(study_data, aes(x = treatment, y = bp_reduction_mmHg, color = treatment)) +
  geom_jitter() +
  theme_minimal() +
  labs(
    title = "Blood Pressure Reduction by Treatment",
    x     = "Treatment",
    y     = "BP Reduction (mmHg)"
    ) +
  theme(legend.position = "none") +
  stat_summary(
      fun = mean,              # compute mean
      geom = "point",          # plot as point
      col = "black",
      size = 3,
      shape = 'square'
  )
```


## Step 4: Write a function

### Number of participants (rows)

```{r}
number_participants <- function(input_data) {
  n_rows <- nrow(input_data) 
  return(n_rows)
}
```

```{r}
number_participants(study_data)
```
### Add number of columns
```{r}
calculate_size <- function(data) {
  n_rows <- nrow(data) 
  n_cols <- ncol(data)
  return(c(n_rows, n_cols))
}
```

```{r}
calculate_size(study_data)
```

### Why functions? To write less code!
E.g., let's look at two different datasets
```{r}
study_data <- read_excel("bp_trial.xlsx")
study_data_2 <- read_excel("bp_trial_v2.xlsx")
```

Without a function:
```{r}
nrow(study_data) 
ncol(study_data)
nrow(study_data_2) 
ncol(study_data_2)
```

With a function we save two lines!:
```{r}
calculate_size(study_data)
calculate_size(study_data_2)
```


